---
title: "01 - Introduction"
author: "Stefano Coretta"
format:
  mono-light-revealjs:
    theme: [default, custom.scss]
    history: false
filters:
  - tachyonsextra
---

```{r}
#| label: setup

library(tidyverse)
theme_set(theme_light())
library(coretta2018itapol)
library(brms)
```


## Statistical modelling

...

## Statistical inference

...

## Two frameworks of inference

...

## Frequentist framework

...

## Bayesian framework

...

## Bayesian regression: basics

::: box-note

- R package brms.

- Same syntax as lme4.

- One function to fit many types of regression models: `brm()`.

:::

## Example data: background

...

## Example data: the data frame

```{r}
#| label: data

library(coretta2018itapol)
data("token_measures")

token_measures
```

## Vowel duration and consonant voicing

```{r}
#| label: vdur-strip
token_measures |> 
  ggplot(aes(c2_phonation, v1_duration)) +
  geom_jitter(width = 0.2, alpha = 0.1) +
  stat_summary(fun.data = "mean_cl_boot", colour = "purple") +
  facet_grid(cols = vars(language))
```


## Vowel duration

```{r}
#| label: vdur-dens
token_measures |> 
ggplot(aes(v2_duration)) +
geom_density(fill = "purple", alpha = 0.5) +
geom_rug(alpha = 0.1)
```

## Gaussian model of vowel duration

$$
\begin{align}
dur & \sim Gaussian(\mu, \sigma)
\end{align}
$$

- $dur$ is vowel duration

- $\sim$ means "is distributed/generated according to".

- $Gaussian(\mu, \sigma)$ a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.

## Fit the model with brms

```{r}
#| label: m-1

m_1 <- brm(
  v2_duration ~ 1,
  family = gaussian,
  data = token_measures,
  cores = 4,
  file = "data/cache/m_1",
)
```

## Model summary

```{r}
#| label: m-1-summary

summary(m_1)
```

## Model summary: 80% CrIs

```{r}
#| label: m-1-summary-0.8
summary(m_1, prob = 0.8)
```

## Plot posterior probability distributions

```{r}
#| label: m-1-plot

plot(m_1, combo = c("dens", "trace"))
```

## Diagnostics: posterior predictive checks

```{r}
#| label: m-1-pp-check
pp_check(m_1, ndraws = 50)
```

## Log-normal model of vowel duration

$$
\begin{align}
dur & \sim LogNormal(\mu, \sigma)
\end{align}
$$

. . .

$$
\begin{align}
log(dur) & \sim Gaussian(\mu_1, \sigma_1)
\end{align}
$$

## Fit the log-normal model

```{r}
#| label: m-2

m_2 <- brm(
  v2_duration ~ 1,
  family = lognormal,
  data = token_measures,
  cores = 4,
  file = "data/cache/m_2",
)
```

## Model summary

```{r}
#| label: m-2-summary

summary(m_2, prob = 0.8)
```

## Posterior predictive checks

```{r}
#| label: m-2-pp-check

pp_check(m_2, ndraws = 50)
```

