---
title: "03 - Categorical predictors"
author: "Stefano Coretta"
format:
  mono-light-revealjs:
    theme: [default, custom.scss]
    history: false
filters:
  - tachyonsextra
execute:
  echo: true
knitr:
  opts_chunk: 
    fig-align: center
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
theme_set(theme_light())
library(coretta2018itapol)
library(brms)
library(htmltools)

library(coretta2018itapol)
data("token_measures")
token_measures <- token_measures |> drop_na(v1_duration)

library(HDInterval)
library(truncdist)
```

## Vowel duration by voicing

```{r}
#| label: vdur-voi
#| echo: false
token_measures |> 
  ggplot(aes(c2_phonation, v1_duration)) +
  geom_violin() +
  geom_jitter(width = 0.1, alpha = 0.2) +
  stat_summary(fun.data = "mean_cl_boot", colour = "red")
```

## Categorical predictors: treatment contrasts (default)

$$
\begin{align}
vdur_i & \sim LogNormal(\mu, \sigma)\\
\mu & = \beta_0 + \beta_1 \cdot \text{voiceless}_i \\
\beta_0 & \sim Gaussian(\mu_0, \sigma_0)\\
\beta_1 & \sim Gaussian(\mu_1, \sigma_1)\\
\sigma & \sim Cauchy_{+}(0, \sigma_2)\\
\end{align}
$$

## Get default priors

```{r}
#| label: get-prior

get_prior(
  v1_duration ~ c2_phonation,
  family = lognormal,
  data = token_measures
)
```

## Categorical predictors: indexing

$$
\begin{align}
vdur_i & \sim LogNormal(\mu_i, \sigma)\\
\mu_i & = \beta_{voicing[i]}\\
\beta_j & \sim Gaussian(\mu_j, \sigma_j) & , \text{for } j = 1..2\\
\sigma & \sim Cauchy_{+}(0, \sigma_3)\\
\end{align}
$$

## Get default priors: indexing

```{r}
#| label: get-prior-2

get_prior(
  v1_duration ~ 0 + c2_phonation,
  family = lognormal,
  data = token_measures
)
```

## Priors

```{r}
#| label: m-4-priors

m_4_priors <- c(
  # Prior for both `b` coefficients
  prior(normal(4.45, 0.55), class = b),
  prior(normal(0, 0.1), class = sigma)
)

m_4_priors
```

## Prior predictive checks

```{r}
#| label: m-4-prior-pp

my_seed <- 9682

m_4_priorpp <- brm(
  v1_duration ~ 0 + c2_phonation,
  data = token_measures,
  family = lognormal,
  prior = m_4_priors,
  sample_prior = "only",
  cores = 4,
  seed = my_seed,
  file = "data/cache/m_4_priorpp"
)
```

## Prior predictive plots

```{r}
#| label: m-4-prior-pp-plot

conditional_effects(m_4_priorpp)
```

## Fit the model

```{r}
#| label: m-4

m_4 <- brm(
  v1_duration ~ 0 + c2_phonation,
  data = token_measures,
  family = lognormal,
  prior = m_4_priors,
  seed = my_seed,
  cores = 4,
  file = "data/cache/m_4"
)
```

## Model summary

```{r}
#| label: m-4-summary

summary(m_4, prob = 0.8)
```

## Posterior probability distributions: outcome (vowel duration)

```{r}
#| label: m-4-cond

conditional_effects(m_4)
```

## Posterior probability distribution: difference of vowel duration

```{r}
#| label: m4-comparisons

library(marginaleffects)

avg_comparisons(m_4, conf_level = 0.8)
```

## Posterior probability distributions

| Posteriors | package         | function                |
|------------|-----------------|-------------------------|
| Outcome    | brms            | `conditional_effects()` |
| Difference | marginaleffects | `avg_comparisons()`     |
